Comparison of Several Stochastic and Deterministic Derivative-free Global Optimization Algorithms

### 2019-04-07 2019/MOTOR_Sovrasov/slides/slides.tex
MOTOR: add initial version of slides

\title{Comparison Of Several Sequential And Parallel Derivative-free Global Optimization Algorithms}
\author{\underline{\textbf{Vladislav~Sovrasov}} and \textbf{Semen~Bevzuk}}
\institute{Lobachevsky State University of Nizhni Novgorod}
\date{}

@@

\begin{frame}
  \frametitle{Globalizer parallel solver}
  \begin{figure}[ht]
    \includegraphics[width=0.95\textwidth]{evolvents_parallel.pdf}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{MIDACO parallel solver}
  The MIDACO solver utilizes reverse communication architecture.
  Reverse communication means that the call of the objective function happens outside and
  independently of the MIDACO source code.
  Relying on this feature the following distributed parallelization scheme was implemented:
  \begin{enumerate}
    \item MIDACO runs at the master node and generates \(n \times p\) new trial points at each
  iteration.
    \item Each of \(n\) distributed computational nodes gets \(p\) points from the master and performs
  \(p\) parallel trials on local computing devices.
    \item Each of \(n\) distributed computational nodes sends \(p\) values of the objective function to
  the master.
  \end{enumerate}

  In this scheme all distributed communications were implemented using the MPI library.
  Parallelism within a single node is powered by the OpenMP standard.
\end{frame}

\begin{frame}
  \frametitle{Cluster environment}
  \begin{center}
  The computational experiments have been carried out on the Lobachevsky supercomputer at  State University of Nizhni Novgorod.
  One node includes 2 Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 GB RAM.

  Each node runs under CentOS 7 Linux with GCC 4.8 compiler and Intel MPI library.
\end{center}
\end{frame}

\begin{frame}
  \frametitle{Results of the parallel algorithms}
  \begin{table}
    \centering
    \caption{Averaged numbers of iterations executed by the parallel algorithms for solving the test
  optimization problems}
    \label{tab:iterations}
    \begin{tabular}{cccccccc}
      \cline{3-8}\noalign{\smallskip}
      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
  \multicolumn{2}{c}{MIDACO}   \\
      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
  \textit{Hard}  \\
      \noalign{\smallskip}\hline
      I & \textbf{1 cluster node}
        & \textit{1} &   25270 (100)  & 55180 (99) & & 27645 (98) & 72068 (71)  \\
      &  & \textit{16} & 1765 (100)  & 3714 (100) & &  1640 (97) & 4304 (70) \\
      \hline \noalign{\smallskip}
  II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
    & \textit{1} & 13056 (100) & 22938 (99) & & 10558 (89) & 27128 (73) \\
  &   & \textit{16} & 732 (100) & 1759 (100)  &  & 1130 (92) & 2254 (73) \\
      \hline \noalign{\smallskip}
  III & \textbf{4 cluster nodes} %\multirow{3}{*}{}
    & \textit{1}  & 5016 (100) & 12703 (100) & & 5777 (94) & 15980 (75) \\
  & & \textit{16} & 367 (100) & 776 (100) & & 529 (88) &  1264 (66) \\
      \hline \noalign{\smallskip}
  VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
    & \textit{1}  & 2103 (100) & 5063 (100) & & 2847 (97) & 9853 (83)\\
  & & \textit{16} & 145 (100)  & 310 (100)  & & 272 (83) & 774 (57)\\
      \hline \noalign{\smallskip}
  V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
    & \textit{1}  & 1155 (100) & 2399 (100) & & 2233 (98) & 6022 (86) \\
  & & \textit{16} & 76 (100)  & 159 (100)  & & 168 (87) & 393 (53)\\
      \noalign{\smallskip}\hline
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}
  \frametitle{Results of parallel algorithms}
  \begin{table}
    \centering
    \caption{Speedup of parallel computations executed by the parallel algorithms}
    \label{tab:speedup}
    \begin{tabular}{cccccccc}
      \cline{3-8}\noalign{\smallskip}
      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{Globalizer} & &
  \multicolumn{2}{c}{MIDACO}   \\
      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
  \textit{Hard}  \\
      \noalign{\smallskip}\hline
  I  & \textbf{1 cluster node}  %\multirow{3}{*}{}
      & \textit{1}   & 25270 (27.3s) & 55180 (61.8s) & & 27645 (32.2s) & 72068 (132.5s)  \\
    &  & \textit{16} & 14.3(11.7) & 14.9(12.6)  & &  16.9(14.4) & 16.7(14.4) \\
    \hline \noalign{\smallskip}
  II  & \textbf{2 cluster nodes}  %\multirow{3}{*}{}
    & \textit{1}      &   1.9(1.9) & 2.4(2.2)  & & 2.6(1.7) & 2.7(2.3) \\
    &   & \textit{16} & 34.5(20.4) & 31.4(18.8) & & 24.5(18.8) & 32.0(29.1) \\
      \noalign{\smallskip}\hline	\noalign{\smallskip}
  III  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
  & \textit{1}      & 5.0(4.6) & 4.3(4.1) & & 4.8(3.9) & 4.5(4.4) \\
  &   & \textit{16} & 68.8(23.7) & 71.2(25.8) & & 52.3(32.9) & 57.0(50.2) \\
    \noalign{\smallskip}\hline	\noalign{\smallskip}
  VI & \textbf{8 cluster nodes} %\multirow{3}{*}{}
  & \textit{1}    & 12.0(8.7) & 10.9(8.3) & & 9.7(8.2)    & 7.3(9.0)    \\
  & & \textit{16} & 174.1(4.7) & 177.8(5.3) & & 101.6(51.6) & 93.1(84.1) \\
      \noalign{\smallskip}\hline
      V & \textbf{12 cluster nodes} %\multirow{3}{*}{}
      & \textit{1}    & 21.9(11.4)  & 23.0(12.8)  & & 12.4(11.1)  & 12.0(14.8)  \\
      & & \textit{16} & 333.5(2.7) & 347.9(3.0) & & 164.5(82.0) & 183.2(129.7) \\
          \noalign{\smallskip}\hline
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}
  \frametitle{Results of parallel algorithms}
  \begin{figure}[ht]
    \centering
    \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{speedup_midaco_globalizer_p_1.pdf}}\label{fig:speedup_p1}}
    \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{speedup_midaco_globalizer_p_16.pdf}}\label{fig:speedup_p16}}
    \caption{Speedup demonstrated by the parallel algorithms when solving problems from the GKLS 4d Hard class}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Results of parallel algorithms}
  \begin{figure}[ht]
    \centering
    \subfloat[$p=1$]{{\includegraphics[width=.5\textwidth]{time_midaco_globalizer_p_1.pdf}}\label{fig:time_p1}}
    \subfloat[$p=16$]{{\includegraphics[width=.5\textwidth]{time_midaco_globalizer_p_16.pdf}}\label{fig:time_p16}}
    \caption{Averaged execution time of the parallel algorithms when solving problems from the GKLS 4d Hard class}
  \end{figure}
\end{frame}

>>>>

\title{Comparison of Several Stochastic and Deterministic Derivative-free Global Optimization Algorithms}
\author{\textbf{Vladislav~Sovrasov}}
\institute{Lobachevsky State University of Nizhni Novgorod}
\date{}

@@
